A global compilation reveals divergent biomass partitioning between plant functional types

### 2014-10-31 manuscript/draftpaper.Rnw
Added supporting information on behaviour of simple two-pool allocation model.

\usepackage{amssymb,amsmath}
\usepackage{palatino}
\usepackage{framed}
\usepackage{parskip}
\usepackage{graphicx}
\usepackage[default,osfigures,scale=0.95]{opensans}
\graphicspath{ {./figures/} }
\usepackage{geometry}
\usepackage{authblk}
\usepackage{gensymb}
\usepackage{lineno}
\geometry{verbose,tmargin=3cm,bmargin=3cm,lmargin=3cm,rmargin=3cm}
\bibliographystyle{plain}

@@

>>>>

\usepackage{amssymb,amsmath}
\usepackage{palatino}
\usepackage{framed}
\usepackage{parskip}
\usepackage{graphicx}
\usepackage[T1]{fontenc}
\usepackage[default,osfigures,scale=0.95]{opensans}
\graphicspath{ {./figures/} }
\usepackage{geometry}
\usepackage{authblk}
\usepackage{gensymb}
\usepackage{lineno}
\geometry{verbose,tmargin=3cm,bmargin=3cm,lmargin=3cm,rmargin=3cm}
\bibliographystyle{plain}

@@

\subsection*{A simple biomass allocation model}

We have shown that biomass partitioning to leaf biomass decreases quickly with increasing plant size. However, this does not yield a model for the allocation of net primary production (NPP) to leaf and stem components, since partitioning depends on allocation as well as turnover. Here we construct a minimal model of biomass allocation to evaluate the fixed allocation model (a model that allocates a fixed proportion of NPP to leaves and stems, not accounting for possible shifts with plant size). We consider two pools: a leaf pool with high turnover, and an aboveground woody pool with slow turnover. Belowground allocation is not considered here, it is implicitly assumed that the allocation to belowground is a fixed proportion of NPP.

Dynamics of leaf mass ($M_F$) and aboveground woody biomass ($M_S$) depend on annual allocation of NPP to these pools ($\phi_F$ and $phi_S$), and turnover rates ($yr^{-1}$) ($r_F$ and $r_S$), as in:

\begin{equation}\label{eq:dmfdt}
\frac{dM_F}{dt} = \phi_F NPP - r_F M_F
\end{equation}

\begin{equation}\label{eq:dmsdt}
\frac{dM_S}{dt} = \phi_S NPP - r_S M_S
\end{equation}
With the constraint that $\phi_F + \phi_S = 1$. We further assume that NPP is proportional to foliage biomass,

\begin{equation}
NPP = a_F M_F
\end{equation}

So that Eqs.~\ref{eq:dmfdt} and~\ref{eq:dmsdt} become,
\begin{equation}
\frac{dM_F}{dt} = (\phi_F  a_F - r_F) M_F
\end{equation}

\begin{equation}
\frac{dM_S}{dt} = (1 - \phi_F)  a_F  M_F - r_S M_S
\end{equation}

These two differential equations may be integrated (using the method of variation of parameters) to yield the solution,

\begin{equation}\label{eq:mfsolution}
M_F(t) = c_2 e^{t(\phi_F a_F - r_F)}
\end{equation}

and

\begin{equation}\label{eq:mssolution}
M_S(t) = \frac{\phi_S a_F c_2}{\phi_F a_F - r_F} \frac{1}{\frac{r_S}{\phi_F a_F - r_F} + 1} e^{t(\phi_F a_F - r_F)}
+ c_4 e^{-r_S t}
\end{equation}

where $c_2 = M_f(0)$ and

$$
c_4 = M_S(0) - \frac{\phi_S a_F c_2}{\phi_F a_F - r_F} \frac{1}{\frac{r_S}{\phi_F a_F - r_F} + 1}
$$

We are here interested in the ratio $M_F/M_S$ and its dependence on plant size. We compared simulations using Eqs.~\ref{eq:mfsolution} and~\ref{eq:mssolution} to the average relationship estimated from our database (see Fig.~\ref{fig:figureSI10}). It is clear from these simulations that $M_F/M_S$ quickly reaches a steady state value, and shows very little decline with increasing plant size. In contrast, the data show a very steep (note the log-log axes) and persistent decline in $M_F/M_S$. These results indicate that fixed allocation model can not adequately be used to model $M_F/M_S$, some of the parameters must be a function of plant size (most likely, the allocation to foliage mass should decline with size).

Finally, it can easily be shown that $M_F/M_S$ converges to the following equation (found by setting $t = \infty$ in Eqs.~\ref{eq:mfsolution} and~\ref{eq:mssolution}, and taking the ratio $M_F/M_S$). To simplify the result, we have also assumed that $r_S = 0$ (because we can assume that $r_S << r_F$).

\begin{equation}
\lim_{t\to \infty} \frac{M_F}{M_S} = \frac{\phi_F - r_F/a_F}{1 - \phi_F|}
\end{equation}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{FigureSI-10_allocationmodel.pdf}
    \caption{caption text}
  \label{fig:figureSI10}
\end{figure}
